---
title: "Citibike"
output: html_document
---

```{r}
library(readr)
library(dplyr)

citibike = read_csv("201606-citibike-tripdata.csv")



```

```{r}
citibike_variables = names(citibike)
citibike_variables
```

Cleaned variable names for citibike
```{r}

citibike_clean = rename(citibike, start_station_id = `start station id`,
                               start_station_name = `start station name`,
                               start_station_latitude = `start station latitude`,
                               start_station_longitude = `start station longitude`,
                               end_station_id = `end station id`,
                               end_station_name = `end station name`,
                               end_station_latitude = `end station latitude`,
                               end_station_longitude =`end station longitude`,
                               birth_year = `birth year`)

```

Get all the unique stations in our data

```{r}

start_stations = citibike_clean %>%
  select(start_station_id, start_station_longitude, start_station_latitude) %>% 
  unique()

end_stations = citibike_clean %>% 
  select(end_station_id, end_station_longitude, end_station_latitude) %>% 
  unique 
```

It looks like all the stations were used as a dropoff and starting point.

Now let's map these stations. To create a map, we load the libary 'maps' and plot the min./max latitude and min/max longitutde.
**IMPORTANT** 
Longitude data is the x-axis
Latitude data is the y-axis
```{r}
library(RgoogleMaps)

#We are giving the map 1 standard deviation bigger than the ends of the locations
long_min <- min(start_stations$start_station_longitude)
long_max <-  max(start_stations$start_station_longitude)

lat_min <- min(start_stations$start_station_latitude)
lat_max <- min(start_stations$start_station_latitude)

lon <- c(long_min, long_max)
lat <- c(lat_min, lat_max)

center <- c(mean(lat), mean(lon))
zoom <- min(MaxZoom(range(lat), range(lon)))

citibikeMap <- GetMap(center=center,
                      zoom=zoom,
                      maptype="terrain",
                      destfile="citimap.png",
                      )

mymarkers <- start_stations %>% 
  select(start_station_latitude, start_station_longitude) %>% 
  rename(lat = start_station_latitude,
         lon = start_station_longitude) %>% 
  mutate(size = "small",
         col = "red",
         char = "") %>% 
  sample_n(100)

citibikeMap2 <- GetMap.bbox(lonR = range(lon),
                            latR = range(lat),
                            center = center,
                            destfile = "citistations.png",
                            markers = mymarkers,
                            zoom = 11,
                            maptype = "terrain")
```

Let's try to map using ggplot2

```{r}

library(rgdal)
library(ggplot2)

counties <- readOGR("nybb_16c/nybb.shp", layer="nybb")


```

```{r}
# map the counties

nymap <- ggplot() +
  geom_polygon(data = counties,
               aes(x=long, y=lat, group=group))

ggplot() +
  geom_point(data = stations,
             aes(x = start_station_longitude,
                 y = start_station_latitude),
             color = "red")
```

Let's combine the 2
```{r}
nymap +
  geom_point(data = stations,
             aes(x = start_station_longitude,
                 y = start_station_latitude))

proj4string(counties)

```

```{r}

class(start_stations)
## [1] "data.frame"
coordinates(start_stations)<-~start_station_longitude+start_station_latitude
class(start_stations)
## [1] "SpatialPointsDataFrame"
## attr(,"package")
## [1] "sp"

# does it have a projection/coordinate system assigned?
proj4string(start_stations) # nope
## [1] NA

# we know that the coordinate system is NAD83 so we can manually
# tell R what the coordinate system is
proj4string(start_stations)<-CRS("+proj=longlat +datum=NAD83")

# now we can use the spTransform function to project. We will project
# the mapdata and for coordinate reference system (CRS) we will
# assign the projection from counties

mapdata<-spTransform(start_stations, CRS(proj4string(counties)))

# double check that they match
identical(proj4string(start_stations),proj4string(counties))
## [1] TRUE

```

```{r}

mapdata <- data.frame(mapdata)

# We;'re not dealing with lat/long but with x/y
# this is not necessary but for clarity change variable names
names(mapdata)[names(mapdata)=="start_station_longitude"]<-"x"
names(mapdata)[names(mapdata)=="start_station_latitude"]<-"y"

# now create the map
ggplot() +geom_polygon(data=counties, aes(x=long, y=lat, group=group))+  geom_point(data=mapdata, aes(x=x, y=y), color="red")
```

```{r}

ggplot() +  
    geom_polygon(data=counties, aes(x=long, y=lat, group=group), fill="grey40", 
        colour="grey90", alpha=1)+
    labs(x="", y="", title="Building Area Within 1000m")+ #labels
    theme(axis.ticks.y = element_blank(),axis.text.y = element_blank(), # get rid of x ticks/text
          axis.ticks.x = element_blank(),axis.text.x = element_blank(), # get rid of y ticks/text
          plot.title = element_text(lineheight=.8, face="bold", vjust=1))+ # make title bold and add space
    geom_point(aes(x=x, y=y), data=mapdata, alpha=1, size=3, color="grey20")+# to get outline
    geom_point(aes(x=x, y=y), data=mapdata, alpha=1, size=2)+
    scale_colour_gradientn("Building\narea (sq-km)", 
        colours=c( "#f9f3c2","#660000"))+ # change color scale
    coord_equal(ratio=1) # square plot to avoid the distortion

```

